AiPoker - Implementation of a Poker Agent

### 2014-11-03 report/paper.tex
Spelling corrections and progress on report

%% Abstract section.
\abstract{ 
Poker is a well known card game worldwide and has become very popular to play online for the past few years. Together with the interest of playing the game online an interest of trying to create the best poker agent for online poker became popular. This report describes a project to investigate the possibilities with creating a Poker Agent with the logic of a Monte Carlo Tree Search in combination a few smaller well known algorithms. The result is a poker agent which plays with realistic reasoning.
}
%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Monte Carlo Tree Search, Texas Hold'em}

@@

%% the only exception to this rule is the \firstsection command
\firstsection{Background}
\maketitle 
Texas Hold'em is a popular variant of poker, not just to play with friends but also online against other people for money. The challenge of creating the best poker agent has attracted a lot of people and is a popular subject in AI development.
\\*\\*
\textbf{Type of Games}\\*
There are many different games today and the strategy to solve these games depends on the type of game. In game theory games can be classified due to the knowledge if the transition to a new state within the game depends on a stochastic process and if there is some hidden information. Poker belongs with the most difficult kind of card game to solve discretely since it is an stochastic game with imperfect information\cite{Kleij}.
\\*\\*
\textbf{Perfect Information vs Imperfect Information}\\*
In perfect information games both players can observe the complete state of the game at any given time, this is the case in games such as Chess and Go. Contrary to this we have imperfect information games where each player is unable to observe the complete state, one example for this is poker were neither player can see the other players pocket cards. 
\\*\\*
\textbf{Deterministic vs Stochastic}\\*
In deterministic games the next state of the game is uniquely determined by an action taken at the current state. In contrast, a stochastic game is a game where the player does not have control over the transition to the next state. For example in poker where the new cards is drawn from a scrambled deck.

@@

The subdivision in (table \ref{tab:table1}) exhibit that games like Chess and Go belongs to the category of games that is less complex to try to solve with an AI agent. Poker belongs to one of the most complex games to try to solve.

\subsection{Related Work}
\subsubsection{Game Theory}
In 1997 Koller and Pfeffer\cite{Koller} made a program called Gala, the program uses strategies from game theory to find optimal strategies in imperfect information games. With the help of Gala they were able to solve several games but they come to the conclusion that the number of states in poker is too large to be solvable with the current hardware. However they suggested a solution that maybe it was possible to map several similar states onto one abstract state resulting in an abstract game tree much smaller than the original. 
It was from this idea that (Billings et al 2003)\cite{billingsburch} successfully abstracted the game down from a search space of $O(10^{18})$ to $O(10^{7})$. They then used Linear optimization to find a Nash equilibrium solution to the game. From this they could take any action, map it to the abstract game tree, find a solution and then map it back to the real game. This resulted in a poker agent (PsOPti) that was far better than anything else at the time.

If PsOpti would play a real nash equilibrium strategy no human player would be able to beat it in the long run. But the problem is that it is just an approximation and there will therefore find ways to exploit it. And as PsOpti ignore any action from the human player it allows the player to systematically search for weaknesses in the without fear of being punished.

\subsubsection{Expert Systems}
Billings et al.(1998)\cite{billingspapp} developed an AI agent called Loki. It is an expert system that uses a set of rules and formulas based on the strength of the cards to determine the next action. It uses no opponent modeling in heads up poker. It can be challenging to define strong expert systems as it can be hard to define rules that covers all aspects of the game. There can also be problems with conflicting rules and missing rules (situations that is not covered by any rules). Another problem is that expert systems require that you have great knowledge about the game as the AI will never play better then the player defining the rules.

\subsubsection{Opponent Modeling}
Papp (1998)\cite{papp} introduced the first tries to extend Loki with opponent modeling. The opponent model was a quite simple and just counted the opponents actions in the different streets (preflop, flop, turn, river), but even this simple opponent modeling increased the strength of the AI.

Davidson et al. (2000)\cite{davidson} tried to improve the opponent modeling in Loki with artificial neural networks (ANNs) to predict the opponent's actions, and achieves about 85\% accuracy at predicting the moves. This is a significant improvement over the previous implementation and the new system was named Poki. The problems with ANNs is that they can be slow to train and thus they will make weak plays against players who frequently change their play style.

\subsubsection{Search Trees}
Billings et al. (2004)\cite{billingsdavidson} was one of the first to use game trees search in a poker AI. Their Ai, Vexbot uses two modified versions of Expectiminimax called Miximax and Miximix that extends Expectiminimax to games with hidden information. Vexbot makes an exhaustive  depth first search of the entire search tree for two player poker.

Van den Broeck et al. (2009)\cite{broeck} were the first to apply Monte Carlo search to game trees for poker agents. They focused on 10 player no-limit poker which have enormous game trees which makes methods such as the one described above impossible to use. Here it is impossible to search the whole game tree which makes sampling parts of the game tree the only option.

@@

One possible use for Monte Carlo methods is to determine the value of an integral as illustrated in fig(\ref{fig:integralex}). By randomly placing dots and label them as under or over the function value we can get an estimate of the area by taking the area of the square encapsulating the function times the ratio between the number of points under the function with the total amount of points. The approximated value given by the Monte Carlo simulation will converge to the correct value as the number of sample points goes to infinity.

The insight that random sampling can be used to solve problems that is too complex to solve analytically can be used to evaluate the best action in huge game trees. For a given node $N$ we can estimate its value by running a number of simulations of how the game may end from that node. The result from those simulations can then be used to give an approximation for the value of that state and as $N$ grows it will converge to the result from a Minimax search.

Monte-Carlo methods for game trees have made huge improvements for game which was previously very hard for AIs. One example is Go in which bots powered with MCTS now can reach the highest level of play.

\section{Theory}
\subsection{Texas Hold'em no Limit}
\textit{(If you already are familiar with the game of Texas Hold'em, you can skip to section 2.2).}
There are many forms of poker, with the most popular being the Texas Hold'em. Texas Hold?em can be played in two different styles, limited and no-limits. In limited poker the amount you bet or raise is fixed, and there is an upper amount of times that you can raise. However in no limited there is no restriction to the amount of money or times you can raise.

\subsubsection{Main Concept}
The game aims to win as many hands as possible and to get the opponent(s) coins and when a player is out of coins he has lost the game. When everyone except one player has run out of coins the game is over and a winner is designated.

Each player gets two cards and the player with the best five combination of the cards on the players hand and the cards on the poker board wins the round.

In order to always have money in the pot of each round and take away the risk of a player to be to defensive and never bet a small fee i circulating around the players through the game called the blind. The blind is two amounts called big and small blind. When the game starts a player gets randomly selected to pay the big blind and the player next on the left to him gets to pay the small blind. When a round is played the blind moves one step to the left so that the person who had the small blind get to pay the big blind this turn and the player left to him gets the small blind. If there are only two players they just switch big and small blind every round. The big blind is the double amount of the small blind and is a pre set sum.

The game consist of five stages where the players can decide how to proceed in the game.
These stages are pre-flop, flop, turn, river and showdown.

@@

\textbf{Pre Flop}\\*
In this stage the players get two cards each from the dealer. The cards are called the pocket cards and the player who is first is able to either bet, call if he has not paid the big blind, check if he has paid the big blind or fold. Since you can not form a five combination of cards yet only a speculation of how good your pocket cards are can be made. If a player bets or raises all other players get the opportunity to either call, raise or fold max three times, after that the players who has not folded moves on to the flop.
\\*\\*
\textbf{Flop}\\*
The flop is when the dealer puts out the three first cards on the table. These cards are visible to all players. The pocket cards can now be combined with the flop cards in order to create a strong hand. The players will make moves in the same procedure as in the pre flop stage.
\\*\\*
\textbf{Turn}\\*
Now the dealer adds a fourth card to the flop. The players will now bet again.
\\*\\*
\textbf{River}\\*
The dealer adds a fifth card to the table. This is the last card which is dealt. After the last betting round, the game moves on to showdown. 
\\*\\*
\textbf{Showdown}\\*
In this phase the final betting is completed and the players who has not folded will now show their cards in order to decide who is the winner of this round. The winner will get the full pot and then a new round will start with the pre flop.
\\*\\*
\textbf{The different five combination of cards}\\*
There are nine different types of combination of cards and here they are listed in descending order:
\begin{itemize}
  \item \textit{Straight Flush}: When the player has five cards in a consecutive order of value within the same suit. If more than one player have a Straight Flush the player with the highest values of his/her cards wins. The best variant is the Royal Straight Flush when the player has Ace, King, Queen, Knight and a ten within the same suit.
  \item \textit{Four of a kind}: It is a hand that contains four of the same value, which means in all the different suits. If different players have the same type of hand the hand with the highest four of a kind wins.
  \item \textit{Full House}: This hand is a combination of three of a kind and a pair. If more than one player have a Full House the player with the highest value three of a kind wins, if two players have the same three of a kind the player with the highest value of one pair wins.
  \item \textit{Flush}: A flush is when a player has five cards with the same suit, if more than one player have a flush the one with the highest card wins. If they have the same highest card then the the second highest card is checked and so on.
  \item \textit{Straight}: When you have a five cards in a consecutive order of value. If more than one player has a Straight then the player with the highest card wins.
  \item \textit{Three of a Kind}: When you have three cards of the same value, if more than one player has a Three of a kind then the player with the highest value of the cards wins.
  \item \textit{Two Pair}: When a player has two pair of cards, if more than one player have a two pair the player with the highest pair has the best hand between them.
  \item \textit{One Pair}: When you have one pair of cards, if more than one player have a one pair the player with the highest pair has the best hand between them.
  \item \textit{High Card}: If the card does not match any of the types above then the card with the highest value is considered. 
\end{itemize}

@@

\subsection{Characteristics of a Poker Player}
Like previously described poker is a stochastic game, that makes it hard to control and to define what a good poker player is. A good poker player does not always have to be the one who is playing optimally. A good poker player is hard to define but some of the main characteristics of a good player is that he can read his opponents strategy or way of playing, know when having a good hand to play with and know when to play the hand\cite{pokerlistings}.

A player who tries to maximize his winnings will have to use bluffing. This applies both to the cases when the player holds good cards, and when he holds bad cards. If a player never bluffs his or her strategy becomes predictable and it becomes obvious that the player has good cards as soon as he or she bets. This will lead to the opponent folding and the size of the winning pot being small. 
In the second case, the player might have bad cards. If the player uses bluffing it is possible for him/her to win games where the opponent has better cards.

\subsection{Game Trees and the Motivation for Monte-Carlo Tree Search}
A game tree is a tree structure that from a given state in the game describes all moves to all possible endings of the game.
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.30]{img/MiniMaxGameTree.png}
    \caption{\label{fig:gametree} Part of a tic-tac-toe - game tree.}
  \end{center}
\end{figure}

The above example shows a part of the game tree for a tic-tac-toe game tree. Every node represent a possible state in the game with the root node being the current state and every edge is the action leading to that state. 
The size for a game tree grows very fast when you increase the possible actions that can be taken in each state. For example the game tree for a empty tic-tac-toe board have $255168$ leaf nodes. Two player Texas Hold'em have a search space size of $O(10^{18})$.

With the help of game trees computers can reason about the consequences of their actions and many of the most used AI algorithms are based on them. One of the most used tree search algorithms is the Minimax algorithm. Minimax assumes that both players play an optimal strategy. This means that the player who tries to find the best move always selects the actions that will give him the highest reward thus we call his nodes MAX-nodes. On the other hand the opponent will always select the actions that will give the player the lowest reward and is thus called his nodes MIN-nodes. The value for each leaf node is determined by a heuristic and the value is back propagated up the tree depending on the parent node. Minimax will search the game tree depth first and the result will be the action which gives the highest reward.

Expectiminimax is the continuation of the Minimax algorithm that allows for stochastic events in the game tree such as drawing cards. Expectiminimax trees add an additional type of nodes, chance nodes. At chance nodes a child $c$ is chosen with some probability $P(c)$. This means that a chance node for drawing a card from a full deck have $52$ children with a uniform probability distribution of $1/52$. The value for a chance node is the weighted average of all its children.

@@

It was this problem that led to the development of Monte-Carlo tree search(MCTS). Monte-Carlo tree search works around the need to search the game tree exhaustively by estimating the best action by sampling the tree in its most promising parts.

\section{Method}
\subsection{Basic Game Engine}
In order to even begin implementing an AI, the core poker game had to be implemented first. This was made by first implementing an agent that played at random. For simplicity the game implemented in this study is a two player game with the limited edition of Texas Hold'em where a human player faces a computer AI. The player who starts is chosen at random when the game is started. The player that starts then puts one dollar in to the pot. The other player puts two dollars in the pot. Then the game is turn based like ordinary Texas Hold'em, the players can bet, check, call, raise and fold. And after each betting round cards are dealt to the table and the players combine the cards on hand with the cards on the table in order to get the strongest hand possible.

\subsection{AI-techniques}
A couple of different AI-techniques were implemented in this study. They are described in this section. Also some tests were the different AIs were competing against each other were made.
\subsubsection{Monte Carlo Tree Search}
MCTS is a best-first search strategy. Every node in the tree contain two values, an estimate of the expected value $V(P)$ of the reward $r(P)$ and the number of times the node have been visited $n_{i}$. MCTS only starts with the root node and incrementally build the tree by repeating the following 4 steps.\\*\\*
\textbf{Selection}
\\*Selection is made by starting from the root node and then recursively select nodes until a leaf node is reached (this does not have to be a leaf of the game tree). To determine which node is the optimal one, the upper confidence bounds (UCB1) formula is used\cite{levente}:

\begin{equation} \label{eq:upperconfidence}
	v_{i} + C \cdot \sqrt{\frac{\ln N}{n_{i}}}
\end{equation}
The first term vi is the estimated value of the node and is responsible for exploitation of good nodes, the second term is made up by the total number of visits of the parent node $N$ and the total number of visits of the current node $n_{i}$. The constant $C$ allows tuning of the exploration-exploitation trade-off. 
\\*\\*
\textbf{Expansion}
\\*If the leaf node reached is not an terminating node, which means that the game is not over at the current node. Then create one or more child nodes from the node currently at. This step is illustrated in fig(\ref{fig:expansion}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree1.png}
    \caption{\label{fig:expansion} Illustration of how the green node is added to the tree.}
  \end{center}
\end{figure}
\\*
\textbf{Simulation}
\\*The third step in MCTS is the simulation step. Here a game is simulated from the state in the new node until a result is achieved. The value of the reached result is recorded.
\\*\\*
\textbf{Backpropagation}
\\*The last step is to update the expected value and the selection counter for all nodes along the explored path by backpropagating the recorded result from the simulation step. This step is illustrated in fig(\ref{fig:backprop}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree2.png}
    \caption{\label{fig:backprop} Illustration of how the path to the node is updated with the score from the simulation in the whole tree.}
  \end{center}
\end{figure}
When the back propagation has been done the whole chain starts over again starting with the selection stage\cite{Chaslot}. The use of the UTC selection method makes the game tree grow asymmetrically as it focus its search in the more promising parts of the tree.

One of the pros with MCTS is that it is anytime, that is it will give a valid solution to the problem even if it is interrupted before it is finished. However the quality of the solution is an estimate and is expected to be better the more time the MCTS keeps running.

\subsubsection{Hand Strength}
Since Monte Carlo is a heuristic search which simulate a set of outcomes in a game to calculate what action to make the logic so far does not consider how good the cards on the agents hand is compared to other hands. To add this logic to the agent an algorithm could be used that would return a probability of how strong the current hand cards is with the cards on the table. This probability of the Hand Strength may, at the earliest be calculated when the flop has been dealt\cite{wenkai}.

The Hand Strength is calculated by comparing the cards on the hand with all possible combination of two cards that the opponent can have and sum the times the agent's cards wins compared to the opponent and divide by the number of combinations that was simulated.

\subsubsection{Hand Potential}
The hand potential algorithm calculates the quality of the hand as the game goes on. The algorithm is similar to the hand strength algorithm. The difference is that the hand potential algorithm considers all possible cards that have not been revealed yet. It is also heavier to calculate since it has to go through every possible combination of cards that could possibly end up on the table. It also takes into account every possible card the opponent might have on hand\cite{wenkai}.

\subsubsection{Effective Hand Strength}
By combining the hand strength with the hand potential a probability of winning can be calculated, and this is called the effective hand strength and is calculated as follows\cite{wenkai}:

\begin{equation} \label{EHS}
	EHS = HS \cdot (1-NPot) \cdot PPOT
\end{equation}

\subsubsection{Bucketing}
Since effective hand strength (EHS) needs at least the flop in order to be used for decision making, the Bucketing algorithm was implemented  as a complement to the EHS algorithm. This algorithm is only used in in the pre-flop phase where no board cards are visible \cite{schweizer}.
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.60]{img/bucketlist.png}
    \caption{\label{fig:bucketlist} The bucket list.}
  \end{center}
\end{figure}

Fig(\ref{fig:bucketlist}) gives every two card combination a value between 0 and 4 where 4 is a good combination and 0 is a bad combination. This list is the source of decision making in the pre-flop phase when using the EHS-algorithm as logic in the poker agent.

\subsubsection{Randomness}
All previously mentioned algorithms take cards as input and gives the probability of the cards in different way to make the best decision with the current hand cards. Since a good poker player needs to be able to bluff, a random function has been implemented in order to give the agent a certain level of unpredictability. The agent will make a random move approximately 5\% of the time.

\subsection{Implementation}
\subsubsection{Main Design}
The application implemented in this study was implemented in Javascript using the library Three.js for visualization. Three.js is a webGL library which provides high level functions for creating a 3D scene.

The application is implemented in an object oriented manner. The main application file is called main.js and it is here the scene is created and where all the other classes needed are initialized. The file main.js also contains all the click listeners for the buttons the human player uses to do moves. An overview over the game architecture is shown in fig(\ref{fig:arcitect}). 
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.50]{img/arcitect.png}
    \caption{\label{fig:arcitect} Illustration of how the path to the node is updated with the score from the simulation in the whole tree.}
  \end{center}
\end{figure}

>>>>

%% Abstract section.
\abstract{ 
Poker is a well known card game worldwide and has become very popular to play online for the past few years. This project aims at investigating the possibilities with creating a Poker Agent using the logic of a Monte Carlo Tree Search in combination with a few smaller well known algorithms. The result is a poker agent which plays with realistic reasoning.
}
%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Monte Carlo Tree Search, Texas Hold'em}

@@

%% the only exception to this rule is the \firstsection command
\firstsection{Background}
\maketitle 
Texas Hold'em is a popular variant of poker, not just to play with friends but also online against other people for money. The challenge of creating the best poker agent has attracted a lot of people and is a popular subject in AI development.
\\*\\*
\textbf{Type of Games}\\*
There are many different types of games and the strategy to solve these games depends on the type of game. In game theory games can be classified due to the knowledge of the transition to a new state within the game depends on a stochastic process, and if there is hidden information. Poker belongs to the most difficult kind of games to solve discretely since it is an stochastic game with imperfect information\cite{Kleij}.
\\*\\*
\textbf{Perfect Information vs Imperfect Information}\\*
In perfect information games both players can observe the complete state of the game at any given time, this is the case in games such as Chess and Go. Contrary to this we have imperfect information games where each player is unable to observe the complete state, one example of this is poker were neither player can see the other player's cards. 
\\*\\*
\textbf{Deterministic vs Stochastic}\\*
In deterministic games the next state of the game is uniquely determined by an action taken at the current state. In contrast, a stochastic game is a game where the player does not have control over the transition to the next state. For example in poker where the new cards are drawn from a scrambled deck.

@@

The classification in (table \ref{tab:table1}) shows that games like Chess and Go belong to the category of games that are less complex to solve with an AI agent. Poker belongs to one of the most complex games to solve, since it is both stochastic and contains imperfect information.

\subsection{Related Work}
\subsubsection{Game Theory}
In 1997 Koller and Pfeffer\cite{Koller} made a program called Gala. Gala uses strategies from game theory to find optimal strategies in imperfect information games. With the help of Gala they were able to solve several games but they came to the conclusion that the number of states in poker is too large to be solvable with the hardware available at the time. However, they suggested that maybe it was possible to map several similar states onto one abstract state resulting in an abstract game tree much smaller than the original. 
It was from this idea that (Billings et al 2003)\cite{billingsburch} successfully abstracted the game down from a search space of $O(10^{18})$ to $O(10^{7})$. They then used Linear optimization to find a Nash equilibrium solution to the game. From this they could take any action, map it to the abstract game tree, find a solution and then map it back to the real game. This resulted in a poker agent (PsOPti) that was far better than anything else at the time.

If PsOpti would play a real Nash equilibrium strategy no human player would be able to beat it in the long run. But the problem is that it is just an approximation and there will therefore be ways to exploit it. As PsOpti ignore any action from the human player it allows the player to systematically search for weaknesses without fear of being punished.

\subsubsection{Expert Systems}
Billings et al.(1998)\cite{billingspapp} developed an AI agent called Loki. It is an expert system that uses a set of rules and formulas based on the strength of the cards to determine the next action. It uses no opponent modeling in heads up poker. It can be challenging to define strong expert systems as it can be hard to define rules that cover all aspects of the game. There can also be problems with conflicting rules and missing rules (situations that are not covered by any rules). Another problem is that expert systems require that you have great knowledge about the game as the AI will never play better than the defined the rules.

\subsubsection{Opponent Modeling}
Papp (1998)\cite{papp} was the first to try to extend Loki with opponent modeling. The opponent model was quite simple and just counted the opponents actions in the different game stages (preflop, flop, turn, river). This simple model increased the strength of the AI.

Davidson et al. (2000)\cite{davidson} tried to improve the opponent modeling in Loki with artificial neural networks (ANNs) to predict the opponent's actions, and achieved about 85\% accuracy when predicting the moves. This is a significant improvement over the previous implementation, and the new system was named Poki. The problems with ANNs is that they can be slow to train, thus they will play weakly against players who frequently change their style of playing.

\subsubsection{Search Trees}
Billings et al. (2004)\cite{billingsdavidson} were some of the first to use game tree search in a poker AI. Their AI, Vexbot uses two modified versions of Expectiminimax called Miximax and Miximix that extends Expectiminimax to games with hidden information. Vexbot makes an exhaustive depth first search of the entire search tree for a two player game of poker.

Van den Broeck et al. (2009)\cite{broeck} were the first to apply Monte Carlo tree search for poker agents. They focused on 10 player no-limit poker. 10 player no-limit poker has enormous game trees which make methods such as Miximax and Miximix impossible to use. Since it is impossible to search the whole game tree, sampling parts of the game tree is the only option.

@@

One possible use for Monte Carlo methods is to determine the value of an integral as illustrated in fig(\ref{fig:integralex}). By randomly placing dots and label them as either under or over the function value an estimate of the area can be achieved. By taking the area of the square encapsulating the function times the ratio between the number of points under the function and the total amount of points. The approximated value given by a Monte Carlo method will converge to the correct value as the number of sample points goes to infinity. [REF!]

The insight that random sampling can be used to solve problems that are too complex to solve analytically, can be used to evaluate the best action in huge game trees. For a given node $N$, the value can be estimated by running a number of simulations of how the game may end from that node. The result from those simulations can then be used to give an approximation for the value of that node. As $N$ grows it will converge to the result from a Minimax search.

Monte Carlo tree search has made huge improvements for games that were previously very hard for AIs. One example is Go in which agents powered by MCTS now can compete with the best players in the world.

\section{Theory}
\subsection{Texas Hold'em no Limit}
\textit{(If you already are familiar with the game of Texas Hold'em, you can skip to section 2.2).}
There are many forms of poker, the most popular being Texas Hold'em. Texas Hold'em can be played in two different ways, limited and no-limit. In limited poker the bet and raise amount are fixed, there is also an upper limit of how many times the players raise. However in no-limit poker there is no limit for how much the players can bet, and there is also no limit for how many times the payers can raise.
\subsubsection{Main Concept}
The aim of the game is to win as much money as possible. When a player is out of money, that player has lost. When everyone except one player has run out of money the game is over and a winner is designated.

Every round each player gets two cards on hand, and five cards are dealt to the table. The player with the best combination of these seven cards wins the round.

In order to make sure for every round that there is money in the pot, there is small entry fee called the blind. The blind also takes away the risk of players being too defensive. 
The blind consists of two amounts called the big and the small blind. When the game starts a player gets randomly selected to pay the big blind and the player to the left of him/her gets to pay the small blind. When a new round begins, the blinds move one step to the left. If there are only two players they just switch big and small blind every round. The big blind is the double amount of the small blind and is a pre set sum.

The game consists of five stages where the players can decide how to proceed in the game.
The stages are called:  pre-flop, flop, turn, river and showdown.

@@

\textbf{Pre Flop}\\*
In this stage the players get two cards each from the dealer. These cards are called the pocket cards and after the cards have been dealt, the first betting round begins. If a player bets or raises all other players get the opportunity to either call, raise or fold max three times, after that the players who have not folded move on to the flop.
\\*\\*
\textbf{Flop}\\*
The flop is when the dealer deals the three first cards on the table. These cards are visible to all players. The pocket cards can now be combined with the flop cards in order to create a full hand. The players will make moves in the same procedure as in the pre flop stage.
\\*\\*
\textbf{Turn}\\*
Now the dealer adds a fourth card to the table. The players will now bet again.
\\*\\*
\textbf{River}\\*
The dealer adds a fifth card to the table. This is the last card which is dealt. After the last betting round, the game moves on to showdown. 
\\*\\*
\textbf{Showdown}\\*
In this phase the final betting is completed and the players who have not folded yet will now show their cards in order to decide who is the winner of this round. The winner will get the full pot and then a new round will start with the pre flop.
\\*\\*
\textbf{The different five combination of cards}\\*
There are nine different types of combination of cards and here they are listed in descending order:
\begin{itemize}
  \item \textit{Straight Flush}: When the player has five cards in a consecutive order of value within the same suit. If more than one player have a Straight Flush the player with the highest values of his/her cards wins. The best variant is the Royal Straight Flush when the player has Ace, King, Queen, Jack and a ten within the same suit.
  \item \textit{Four of a kind}: It is a hand that contains four of the same value in all the different suits. If more players have the same type of hand the hand with the highest four of a kind wins.
  \item \textit{Full House}: This hand is a combination of three of a kind and a pair. If more than one player have a Full House the player with the highest three of a kind wins. If more players have the same three of a kind, the player with the highest pair wins.
  \item \textit{Flush}: A flush is when a player has five cards within the same suit, if more than one player have a flush the one with the highest card wins. If they have the same highest card then the the second highest card is checked and so on.
  \item \textit{Straight}: When you have five cards in a consecutive order of value in different suits. If more than one player has a Straight then the player with the highest card wins.
  \item \textit{Three of a Kind}: When you have three cards of the same value, if more than one player has a Three of a kind then the player with the highest value of the cards wins.
  \item \textit{Two Pair}: When a player has two pair of cards, if more than one player have a two pair the player with the highest pair has the best hand between them.
  \item \textit{One Pair}: When you have one pair of cards, if more than one player have a one pair the player with the highest pair has the best hand between them.
  \item \textit{High Card}: If the card does not match any of the types above then the card with the highest value is considered. 
\end{itemize}

@@

\subsection{Characteristics of a Poker Player}
Like previously described poker is a stochastic game. That makes it hard to define what a good poker player is. A good poker player does not always have to be the one who is playing optimally. Some of the main characteristics of a good poker player is that he/she can read his/her opponents strategy or way of playing, has knowledge about when having a good hand to play and know when to play that hand\cite{pokerlistings}.

A player who tries to maximize his/her winnings will have to use bluffing. This applies to both of the cases when the player holds good cards, and when the player holds bad cards. If a player never bluffs his or her strategy becomes predictable and it becomes obvious that the player has good cards as soon as he or she bets. This will lead to the opponent folding and the size of the winning pot being small. 
In the second case, the player might have bad cards. If the player uses bluffing it is possible for him/her to win games where the opponent has better cards.

\subsection{Game Trees and the Motivation for Monte-Carlo Tree Search}
A game tree is a tree structure that from a given state in the game describes all moves to all possible endings of the game.
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.30]{img/MiniMaxGameTree.png}
    \caption{\label{fig:gametree} Part of a tic-tac-toe game tree.}
  \end{center}
\end{figure}

The above example shows a part of a tic-tac-toe game tree. Every node represents a possible state in the game with the root node being the current state and every edge is the action leading to that state. 
The size for a game tree grows very fast when the possible actions that can be taken in each state increase. For example the game tree for an empty tic-tac-toe board has $255168$ leaf nodes. Two player Texas Hold'em has a search space size of $O(10^{18})$. [REF!]

With the help of game trees computers can reason about the consequences of their actions and many of the most used AI algorithms are based on them. One of the most used tree search algorithms is the Minimax algorithm. Minimax assumes that both players play an optimal strategy. This means that the player who tries to find the best move always selects the actions that will give that player the highest reward thus these nodes are call MAX-nodes. On the other hand the opponent will always select the actions that will give the player the lowest reward and these nodes are called MIN-nodes. The value for each leaf node is determined by a heuristic and the value is back propagated up the tree depending on the parent node. Minimax will search the game tree depth first and the result will be the action which gives the highest reward.

Expectiminimax is the continuation of the Minimax algorithm that allows for stochastic events in the game tree such as drawing cards in a game of poker. Expectiminimax trees add an additional type of nodes, chance nodes. At chance nodes a child $c$ is chosen with some probability $P(c)$. This means that a chance node for drawing a card from a full deck has $52$ children with a uniform probability distribution of $1/52$. The value for a chance node is the weighted average of all its children.

@@

It was this problem that led to the development of Monte Carlo tree search (MCTS). Monte Carlo tree search works around the need to search the game tree exhaustively by estimating the best action by sampling the tree in its most promising parts.

\section{Method}
\subsection{Basic Game Engine}
In order to even begin implementing an AI, the core poker game had to be implemented first. This was made by first implementing an agent that played at random. For simplicity the game implemented in this study is a two player game with the limited edition of Texas Hold'em where a human player faces a computer AI.

\subsection{AI-techniques}
A couple of different AI-techniques were implemented in this study. They are described in this section. Also some tests where the different AIs compete against each other were made. The results from these tests are presented in the Results section.

\subsubsection{Monte Carlo Tree Search}
MCTS is a best-first search strategy. Every node in the tree contain two values, an estimate of the expected value $V(P)$ of the reward $r(P)$ and the number of times the node has been visited $n_{i}$. MCTS only starts with the root node and incrementally builds the tree by repeating the following 4 steps:\\*\\*
\textbf{Selection}
\\*Selection is made by starting from the root node and then recursively select nodes until a leaf node is reached (this does not have to be a leaf of the game tree). To determine which node is the optimal one, the upper confidence bounds formula (eq(\ref{eq:upperconfidence})) is used. \cite{levente}:

\begin{equation} \label{eq:upperconfidence}
	v_{i} + C \cdot \sqrt{\frac{\ln N}{n_{i}}}
\end{equation}
The first term $v_{i}$ is the estimated value of the node and is responsible for exploitation of good nodes, the second term is made up by the total number of visits of the parent node $N$ and the total number of visits of the current node $n_{i}$. The constant $C$ allows tuning of the exploration-exploitation trade-off. 
\\*\\*
\textbf{Expansion}
\\*If the leaf node reached is not a terminating node, which means that the game is not over at the current node. Then create one or more child nodes from the node currently at. This step is illustrated in fig(\ref{fig:expansion}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree1.png}
    \caption{\label{fig:expansion} Illustration of how the green node is added to the tree.}
  \end{center}
\end{figure}
\\*\\*
\textbf{Simulation}
\\*The third step in MCTS is the simulation step. Here a game is simulated from the state in the new node until a result is achieved. The value of the reached result is recorded.
\\*\\*
\textbf{Backpropagation}
\\*The last step is to update the expected value and the selection counter for all nodes along the explored path by backpropagating the recorded result from the simulation step. This step is illustrated in fig(\ref{fig:backprop}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree2.png}
    \caption{\label{fig:backprop} Illustration of how the path to the node is updated with the score from the simulation in the whole tree.}
  \end{center}
\end{figure}
When the back propagation has been done the whole chain starts over again starting with the selection stage\cite{Chaslot}. The use of the upper bounds confidence selection method makes the game tree grow asymmetrically as it focus its search in the more promising parts of the tree.

One of the advantages of MCTS is that it can be interrupted at any time but still give a valid result. However the quality of the result will increase the longer time MCTS keeps running.

\subsubsection{Hand Strength}
MCTS simulates a set of outcomes in a game to calculate what action to make but it does not consider the strength of the cards on hand. To add this logic to the agent the Hand Strength algorithm was implemented. Hand Strength returns a probability of how strong the current cards on hand are together with the cards on the table, compared to all possible cards that the opponent might have on hand. This probability of the Hand Strength may, at the earliest be calculated when the flop has been dealt\cite{wenkai}.

The Hand Strength is calculated by comparing the cards on the hand with all possible combinations of two cards that the opponent can have and sum the times the agent's cards wins compared to the opponent and divide by the number of combinations that was simulated.

\subsubsection{Hand Potential}
The hand potential algorithm calculates the strength of the cards on hand as the game goes on. The algorithm is similar to the hand strength algorithm. The difference is that the Hand Potential algorithm considers all possible cards that have not been revealed yet. It is also heavier to calculate since it has to go through every possible combination of cards that could possibly end up on the table whereas the Hand Strength algorithm only looks at the revealed cards. It also takes into account every possible card the opponent might have on hand\cite{wenkai}.

\subsubsection{Effective Hand Strength}
By combining the hand strength with the hand potential a probability of winning can be calculated. This is called the effective hand strength and is calculated as follows\cite{wenkai}:

\begin{equation} \label{EHS}
	EHS = HS \cdot (1-NPot) \cdot PPot
\end{equation}
Where $HS$ is the value obtained from the Hand Strength algorithm, $NPot$ is the negative potential from the Hand Potential and $PPot$ is the positive potential from the Hand Potential.

\subsubsection{Bucketing}
Since effective hand strength (EHS) needs at least the flop to be revealed in order to be used for decision making, the Bucketing algorithm was implemented  as a complement to the EHS algorithm. This algorithm is only used in in the pre-flop phase where no cards are visible on the table \cite{schweizer}.
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.60]{img/bucketlist.png}
    \caption{\label{fig:bucketlist} The bucket list.}
  \end{center}
\end{figure}
\\*The table in fig(\ref{fig:bucketlist}) gives every two card combination a value between 0 and 4 where 4 is a good combination and 0 is a bad combination. This list is the source of decision making in the pre-flop phase when using the EHS-algorithm as logic in the poker agent.

\subsubsection{Randomness}
All previously mentioned algorithms take cards as input and gives the probability of the cards in different ways to make the best decision with the current cards on hand. Since a good poker player needs to be able to bluff, a random function has been implemented in order to give the agent a certain level of unpredictability. The agent will make a random move approximately 5\% of the time.

\subsection{Implementation}
\subsubsection{Main Design}
The application was implemented in Javascript using the library Three.js for visualizing the poker game. Three.js is a webGL library which provides high level functions for creating a 3D scene.

The application is implemented in an object oriented manner. The main application file is called main.js and it is here the scene is created and where all the other classes needed are initialized. The file main.js also contains all the click listeners for the buttons the human player uses to do moves. An overview over the game architecture is shown in fig(\ref{fig:arcitect}). 
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.50]{img/arcitect.png}
    \caption{\label{fig:arcitect} A brief overview of the application architecture.}
  \end{center}
\end{figure}
